---
title: "seabirds_final"
output: html_notebook
---

```{r}
library(tidyverse)
library(janitor)
library(readxl)
library(here)
```

## Reading in the data
```{r}
clean_data <- read_csv(here("clean_data/seabirds_cleaned_data.csv"))
clean_data
```


## Selecting the relevant data
```{r}
seabirds <- clean_data %>% 
              select(record_id, common_name, scientific_name, 
                     species_abbreviation, total_sighting, num_group_sighting, 
                     lat)
seabirds
```


## Qustions

### Q1 Which bird had the most individual sightings?
```{r}
seabirds %>% 
  group_by(common_name) %>% 
  summarise(count = sum(n())) %>% 
  arrange(desc(count))

# Wandering albatross	11293
```



### Ext Q1 combining all the albatross as one type
```{r}
seabirds %>% 
  group_by(common_name) %>%
  filter(str_detect(common_name, "(?i)albatross")) %>% 
  ungroup() %>% 
  summarise(count = sum(total_sighting, na.rm = TRUE))

# All albatross 30424	
```



### Q2 Which bird had the highest total count?
```{r}
seabirds %>% 
  group_by(common_name) %>%
  summarise(count = sum(total_sighting, na.rm = TRUE)) %>% 
  arrange(desc(count))

# Short-tailed shearwater	982553	
```



### Ext Q2 combining all the shearwater as one type
```{r}
seabirds %>% 
  group_by(common_name) %>%
  filter(str_detect(common_name, "(?i)shearwater")) %>% 
  ungroup() %>% 
  summarise(count = sum(total_sighting, na.rm = TRUE))

# All shearwater 1394468		
```



### Q3 Which bird had the highest total count above a latitude of -30?
```{r}
seabirds %>% 
  group_by(common_name) %>%
  filter(lat > -30) %>%
  summarise(count = sum(total_sighting, na.rm = TRUE)) %>%
  arrange(desc(count))
  

# Taking above a latitude of -30 as nearer the equator ie 0
# Wedge-tailed shearwater	855	
```



### Ext Q3 combining all the shearwater as one type
```{r}
seabirds %>% 
  group_by(common_name) %>%
  filter(str_detect(common_name, "(?i)shearwater")) %>% 
  filter(lat > -30) %>%
  ungroup() %>% 
  summarise(count = sum(total_sighting, na.rm = TRUE))

# All shearwater seen above a latitude of -30 (nearer the the equator ie 0) 888		
```



### Q4 How many different types of birds were only ever seen in groups of 1?
```{r}
seabirds %>% 
  group_by(common_name) %>% 
  mutate(max_sighting = max(num_group_sighting, na.rm = TRUE)) %>% 
  filter(max_sighting == 1) %>%
  distinct(common_name) 

# 23 birds only seen once
```


### Ext Q4 If by types of birds you mean more individual species with no variations
```{r}
seabirds %>% 
  group_by(common_name) %>% 
  mutate(common_name = if_else(str_detect(common_name, 
                                          "(?i)shearwater"),"shearwater", 
                                                            common_name),
         common_name = if_else(str_detect(common_name, 
                                          "(?i)albatross"), "albatross",
                                                            common_name),
         common_name = if_else(str_detect(common_name, 
                                          "(?i)mollymawk"), "mollymawk",
                                                            common_name),
         common_name = if_else(str_detect(common_name, 
                                          "(?i)petrel"), "petrel",
                                                            common_name),
         common_name = if_else(str_detect(common_name, 
                                          "(?i)prion"), "prion",
                                                            common_name),
         common_name = if_else(str_detect(common_name, 
                                          "(?i)skua"), "skua",
                                                            common_name)
         ) %>% 
    mutate(max_sighting = max(num_group_sighting, na.rm = TRUE)) %>% 
  filter(max_sighting == 1) %>%
  distinct(common_name) 


# 2 individual species of birds only seen once, 
# when the variations are combined as one group
```




### Q5 How many penguins were seen? (Hint: there are many types of penguin)
```{r}
seabirds %>% 
  group_by(common_name) %>%
  filter(str_detect(common_name, "(?i)penguin")) %>% 
  ungroup() %>% 
  summarise(count = sum(total_sighting, na.rm = TRUE))

# 158	penguins 
```



















