---
title: "R Notebook"
output: html_notebook
---



```{r}
library(tidyverse)
library(here)
```


```{r}
candy <- read_csv(here("clean_data/combined_cleaned_data.csv"))
```

```{r}
head(candy)
tail(candy)
```


## Q1 What is the total number of candy ratings given across the three years. (Number of candy ratings, not the number of raters. Don’t count missing values)
```{r}
candy %>% 
  group_by(sweets, rating) %>% 
  filter(rating != is.na(rating)) %>% 
  summarise(count = n()) %>% 
  ungroup() %>% 
  summarise(total_rating_num = sum(count))

# 695037 vote over the three years
```

## Q2 What was the average age of people who are going out trick or treating?
```{r}
candy %>% 
  group_by(year_id, age) %>% 
  filter(trick_or_treating == "Yes") %>% 
  summarise(count = n()) %>% 
  #group_by(year_id) %>% 
  ungroup() %>% 
  summarise(avg_age = mean(age)) %>% 
  floor()

# 35 average age going trick or treating
```


## Q3 What was the average age of people who are not going trick or treating?
```{r}
candy %>% 
  group_by(year_id, age) %>% 
  filter(trick_or_treating == "No") %>% 
  summarise(count = n()) %>% 
  #group_by(year_id) %>% 
  ungroup() %>% 
  summarise(avg_age = mean(age))%>% 
  floor()

# 39 average age not going trick or treating
```


## Q4 For each of joy, despair and meh, which candy bar revived the most of these ratings?

### JOY
```{r}
candy %>% 
  group_by(sweets, rating) %>% 
  filter(rating != is.na(rating),
         rating == "JOY") %>% 
  summarise(count_joy = n()) %>% 
  arrange(desc(count_joy)) %>% 
  head(1)

# any_full_sized_candy_bar with JOY = 7249		
```
### DESPAIR
```{r}
candy %>% 
  group_by(sweets, rating) %>% 
  filter(rating != is.na(rating),
         rating == "DESPAIR") %>% 
  summarise(count_despair = n()) %>% 
  arrange(desc(count_despair)) %>% 
  head(1)

# broken_glow_stick	with DESPAIR = 7540	
```
### MEH
```{r}
candy %>% 
  group_by(sweets, rating) %>% 
  filter(rating != is.na(rating),
         rating == "MEH") %>% 
  summarise(count_meh = n()) %>% 
  arrange(desc(count_meh)) %>% 
  head(1)

# lollipops	with MEH = 1508		
```

### NA
```{r}
candy %>% 
  group_by(sweets, rating) %>% 
  filter(is.na(rating)) %>% 
  summarise(count_na = n()) %>% 
  arrange(desc(count_na)) %>% 
  head(1)

# most skipped question maynards NA = 2349		
```


## Q5 How many people rated Starburst as despair?
```{r}
candy %>% 
  group_by(sweets, rating) %>% 
  filter(sweets == "starburst" &
         rating == "DESPAIR") %>%
  summarise(count = n()) 

# 1865 people hate starburst

```


## Q6 What was the most popular candy bar by this rating system for each gender in the dataset ?

### Male popular candy
```{r}
joy_male <- candy %>% 
                group_by(sweets, rating, gender) %>% 
                filter(gender == "Male", 
                       rating != is.na(rating),
                       rating == "JOY") %>% 
                summarise(count_joy = n()) %>% 
                arrange(desc(count_joy)) 

despair_male <- candy %>% 
                group_by(sweets, rating, gender) %>% 
                filter(gender == "Male", 
                       rating != is.na(rating),
                       rating == "DESPAIR") %>% 
                summarise(count_despair = n()) %>% 
                arrange(desc(count_despair)) 


male_joy_despair <- full_join(joy_male, despair_male, by="sweets")

male_joy_despair %>% 
  group_by(sweets) %>% 
  summarise(diff_j_d = (count_joy - count_despair)) %>% 
  arrange(desc(diff_j_d)) %>% 
  head(1)

# any_full_sized_candy_bar by 1533 votes
```

### Female popular candy
```{r}
joy_female <- candy %>% 
                group_by(sweets, rating, gender) %>% 
                filter(gender == "Female", 
                       rating != is.na(rating),
                       rating == "JOY") %>% 
                summarise(count_joy = n()) %>% 
                arrange(desc(count_joy)) 

despair_female <- candy %>% 
                group_by(sweets, rating, gender) %>% 
                filter(gender == "Female", 
                       rating != is.na(rating),
                       rating == "DESPAIR") %>% 
                summarise(count_despair = n()) %>% 
                arrange(desc(count_despair))


female_joy_despair <- full_join(joy_female, despair_female, by="sweets")

female_joy_despair %>% 
  group_by(sweets) %>% 
  summarise(diff_j_d = (count_joy - count_despair)) %>% 
  arrange(desc(diff_j_d)) %>% 
  head(1)

# any_full_sized_candy_bar by 855	votes
```


## Q7 What was the most popular candy bar in each year?

### 2015
```{r}
joy_2015 <- candy %>% 
                group_by(sweets, rating) %>% 
                filter(str_detect(year_id, "2015"), 
                       rating != is.na(rating),
                       rating == "JOY") %>% 
                summarise(count_joy = n()) %>% 
                arrange(desc(count_joy))

despair_2015 <- candy %>% 
                group_by(sweets, rating) %>% 
                filter(str_detect(year_id, "2015"), 
                       rating != is.na(rating),
                       rating == "DESPAIR") %>%
                summarise(count_despair = n()) %>% 
                arrange(desc(count_despair))


joy_despair_2015 <- full_join(joy_2015, despair_2015 , by="sweets")


joy_despair_2015 %>% 
  group_by(sweets) %>% 
  summarise(diff_j_d = (count_joy - count_despair)) %>% 
  arrange(desc(diff_j_d)) %>% 
  head(1)

# any_full_sized_candy_bar by 4396 votes

```

### 2016
```{r}
joy_2016 <- candy %>% 
                group_by(sweets, rating) %>% 
                filter(str_detect(year_id, "2016"), 
                       rating != is.na(rating),
                       rating == "JOY") %>% 
                summarise(count_joy = n()) %>% 
                arrange(desc(count_joy))

despair_2016 <- candy %>% 
                group_by(sweets, rating) %>% 
                filter(str_detect(year_id, "2016"), 
                       rating != is.na(rating),
                       rating == "DESPAIR") %>%
                summarise(count_despair = n()) %>% 
                arrange(desc(count_despair))


joy_despair_2016 <- full_join(joy_2016, despair_2016 , by="sweets")


joy_despair_2016 %>% 
  group_by(sweets) %>% 
  summarise(diff_j_d = (count_joy - count_despair)) %>% 
  arrange(desc(diff_j_d)) %>% 
  head(1)

# any_full_sized_candy_bar by 983 votes
```

### 2017
```{r}
joy_2017 <- candy %>% 
                group_by(sweets, rating) %>% 
                filter(str_detect(year_id, "2017"), 
                       rating != is.na(rating),
                       rating == "JOY") %>% 
                summarise(count_joy = n()) %>% 
                arrange(desc(count_joy))

despair_2017 <- candy %>% 
                group_by(sweets, rating) %>% 
                filter(str_detect(year_id, "2017"), 
                       rating != is.na(rating),
                       rating == "DESPAIR") %>%
                summarise(count_despair = n()) %>% 
                arrange(desc(count_despair))


joy_despair_2017 <- full_join(joy_2017, despair_2017 , by="sweets")


joy_despair_2017 %>% 
  group_by(sweets) %>% 
  summarise(diff_j_d = (count_joy - count_despair)) %>% 
  arrange(desc(diff_j_d)) %>% 
  head(1)

# any_full_sized_candy_bar by 1499 votes
```


## Q8 What was the most popular candy bar by this rating for people in US, Canada, UK, and all other countries?

### US
```{r}
joy_us <- candy %>% 
                group_by(sweets, rating) %>% 
                filter(country == "US", 
                       rating != is.na(rating),
                       rating == "JOY") %>% 
                summarise(count_joy = n()) %>% 
                arrange(desc(count_joy))

despair_us <- candy %>% 
                group_by(sweets, rating) %>% 
                filter(country == "US", 
                       rating != is.na(rating),
                       rating == "DESPAIR") %>%
                summarise(count_despair = n()) %>% 
                arrange(desc(count_despair))


joy_despair_us <- full_join(joy_us, despair_us , by="sweets")


joy_despair_us %>% 
  group_by(sweets) %>% 
  summarise(diff_j_d = (count_joy - count_despair)) %>% 
  arrange(desc(diff_j_d)) %>% 
  head(1)

# any_full_sized_candy_bar by 2124 votes
```


### UK
```{r}
joy_uk <- candy %>% 
                group_by(sweets, rating) %>% 
                filter(country == "UK", 
                       rating != is.na(rating),
                       rating == "JOY") %>% 
                summarise(count_joy = n()) %>% 
                arrange(desc(count_joy))

despair_uk <- candy %>% 
                group_by(sweets, rating) %>% 
                filter(country == "UK", 
                       rating != is.na(rating),
                       rating == "DESPAIR") %>%
                summarise(count_despair = n()) %>% 
                arrange(desc(count_despair))


joy_despair_uk <- full_join(joy_uk, despair_uk , by="sweets")


joy_despair_uk %>% 
  group_by(sweets) %>% 
  summarise(diff_j_d = (count_joy - count_despair)) %>% 
  arrange(desc(diff_j_d)) %>% 
  head(1)

# cash_or_other_forms_of_legal_tender by 34 votes
```


### Canada
```{r}
joy_can <- candy %>% 
                group_by(sweets, rating) %>% 
                filter(country == "Canada", 
                       rating != is.na(rating),
                       rating == "JOY") %>% 
                summarise(count_joy = n()) %>% 
                arrange(desc(count_joy))

despair_can <- candy %>% 
                group_by(sweets, rating) %>% 
                filter(country == "Canada", 
                       rating != is.na(rating),
                       rating == "DESPAIR") %>%
                summarise(count_despair = n()) %>% 
                arrange(desc(count_despair))


joy_despair_can <- full_join(joy_can, despair_can , by="sweets")


joy_despair_can %>% 
  group_by(sweets) %>% 
  summarise(diff_j_d = (count_joy - count_despair)) %>% 
  arrange(desc(diff_j_d)) %>% 
  head(1)

# any_full_sized_candy_bar by 247 votes
```


